Explore normal cell types in Neuroblastoma samples

### 2025-06-09 analyses/infercnv-consensus-cell-type/exploratory-notebooks/SCPCP000004/01_nb-consensus-cell-types.Rmd
Apply suggestions from code review
Co-authored-by: Ally Hawkins <54039191+allyhawkins@users.noreply.github.com>

We should remove any cell groups which might correspond to tumor cells, including `stromal`, `neuronal`, and `stem cell`.
The `NA` values here correspond to `Unknown` cells, which we'll also remove.
This means we'll keep groups `endothelial`, `epithelial`, and `immune`.

```{r}
exclude_groups <- c("stromal", "neuronal", "stem cell")
celltype_df <- celltype_df |>
  dplyr::filter(
    !(reference_cell_group %in% exclude_groups),
    consensus_annotation != "Unknown"
  )
```

Let's begin by counting how many cells in each cell group we have per library

@@

We'll look at this data in barplot form to see, per-library, the relative proportions of each normal cell type group out of all normal cells in the library.

```{r}
celltype_prop_df <- celltype_count_wide_df |>
  # long again, but keep all_normal column for calculating proportion
  tidyr::pivot_longer(
    -c(library_id, all_normal),
    names_to = "reference_cell_group",
    values_to = "n"
  ) |>
  dplyr::mutate(cell_group_prop = n / all_normal)

ggplot(celltype_prop_df) +
  aes(x = library_id, y = cell_group_prop, fill = reference_cell_group) +
  geom_col() +
  labs(y = "Proportion of normal cells in group") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
  )
```

@@

* All normal cells
* ALl immune cells
* All endothelial cells

@@

```{r}
# update the count data frame without categories we don't need
celltype_count_df <- celltype_count_wide_df |>
  dplyr::select(-adipocyte, -epithelial) |>
  tidyr::pivot_longer(
    -library_id,
    names_to = "reference_cell_group",
    values_to = "n"
  )

>>>>

We should remove any cell groups which might correspond to tumor cells, including `stromal`, `neuronal`, and `stem cell`.
The `NA` values here correspond to `Unknown` cells, which we'll also remove.
This means we'll keep groups `endothelial`, `epithelial`, `adipocyte`, and `immune`.

```{r}
# First, we want to count the total number of cells to use for later plotting
total_cells_df <- celltype_df |>
  dplyr::count(library_id, name = "total_cells")

# Now we can remove unwanted cells
exclude_groups <- c("stromal", "neuronal", "stem cell")
celltype_df <- celltype_df |>
  dplyr::filter(
    !(reference_cell_group %in% exclude_groups),
    consensus_annotation != "Unknown"
  )
```

Let's begin by counting how many cells in each cell group we have per library.

@@

We'll look at this data in barplot form to see, per-library, the relative proportions of each normal cell type group out of _all cells_ in the library.
The height of each bar is the total fraction of normal cells in that library to begin with; note the Y-axis only ranges from around `[0,0.3]`.

```{r}
celltype_prop_df <- celltype_count_wide_df |>
  dplyr::select(-all_normal) |>
  # long again, but keep all_normal column for calculating proportion
  tidyr::pivot_longer(
    -library_id,
    names_to = "reference_cell_group",
    values_to = "n"
  ) |>
  dplyr::left_join(total_cells_df) |>
  dplyr::mutate(cell_group_prop = n / total_cells)

ggplot(celltype_prop_df) +
  aes(x = library_id, y = cell_group_prop, fill = reference_cell_group) +
  geom_col() +
  labs(y = "Proportion of cells in normal group") +
  theme(
    legend.position = "bottom",
    axis.text.x = element_text(angle = 30, hjust = 1, size = 7)
  )
```

@@

* All normal cells
* All immune cells
* All endothelial cells

@@

```{r}
# update the count data frame without categories we don't need.
# note that this does not remove adipocyte or epithelial counts from the `all_normal` group; it just
# removes them as standalone categories
celltype_count_df <- celltype_count_wide_df |>
  dplyr::select(-adipocyte, -epithelial) |>
  tidyr::pivot_longer(
    -library_id,
    names_to = "reference_cell_group",
    values_to = "n"
  )
